Forecasting Influenza Hospitalizations Using a Bayesian Hierarchical Nonlinear Model with Discrepancy

### 2025-08-28 response.tex

\begin{enumerate}[1.]
\item The entire Bayesian inference scheme seems somewhat heuristic.

\begin{enumerate}[a.]
\item The ILI model is run separately and then feeds into the hospitalization model. So there is no
flow of information from the hospitalization data to the ILI model, which there would be in a fully
Bayesian procedure. Is this done on purpose and if so what is the motivation? The proposed
procedure reminds me of the "Bayesian cut" in the version described by Plummer
\href{(https://link.springer.com/article/10.1007/s11222-014-9503-z)}{(https://link.springer.com/article/10.1007/s11222-014-9503-z)}.
This may be a useful reference.

\spencer{We appreciated this comment and the included reference to the Bayesian
cut. This led to several new insights. 
We added a couple paragraphs of discussion to address this. They appear in
the revised manuscript as below.

``This two-component modeling is remeniscent of the "Bayesian cut" as 
described in [44] and [41]. Cutting a model
into two component models or ``modules`` may be justfied for several reasons
including to reduce time to fit the model, avoiding mixing issues of posterior
sampling distributions, improving predictive performance or to prevent
model misspecification in one model to influence the other 
[41, 27, 26, 44].

The decision to cut the forecast model into two component models
was made for several reasons.
When fitting the fully Bayesian joint model, the time to fit a single model using
Markov chain Monte Carlo (MCMC) sampling and
obtain forecasts was sometimes many hours whereas fitting the two models
separately took a fraction of the time. When assessing convergence of model
parameters, it was rare that posterior sampling chains for the joint model
showed convergence even after 55,000 iterations, whereas the component model 
fits gave no signs which raised concern for convergence. And most importantly,
the forecasts from the component models generally outperformed those of the 
joint model in terms of minimizing proper scoring rules.``}

@@

% \spencer{In the forecasting competition, the data comes generally by the afternoon and
% forecasts are due by the evening. We initially did a full Bayesian procedure
% (minus the MLE selection of the $\lambda$ parameter in the ASG ILI model), but fit 
% was very slow and made it difficutl to get forecasts submitted in time. We switched
% to using two separate models because of the time saved in model fitting. We also
% found that in terms of the proper scoring rules and predictive interval coverage,
% there wasn't a drop in performance although the parameter estimation did 
% change.
% 
% \begin{equation}
% \begin{aligned}
%     p(\tilde{H}_{s,w^*} | \textbf{ILI}, \textbf{H}) &= \int 
%        p(\tilde{H}_{s,w^*} | \textbf{ILI}, \textbf{H}, \boldsymbol{\alpha}) 
%        p(\boldsymbol{\alpha} | \textbf{ILI}, \textbf{H}) d\boldsymbol{\alpha} \\
%     p(\tilde{H}_{s,w^*} | \textbf{ILI}, \textbf{H}, \boldsymbol{\alpha}) &= 
%        \int p(\tilde{H}_{s,w^*} | \textbf{H}, \textbf{ILI}, 
%        \widetilde{ILI}_{s,w^*}, \boldsymbol{\alpha}) 
%        p(\widetilde{ILI}_{s,w^*} | \textbf{ILI}) d\widetilde{ILI}_{s,w^*}
% \end{aligned}
% \end{equation}
% }

\item The authors employ a "modular Bayesian approach" mixing maximum likelihood and
Bayesian estimation. The authors should provide more ample references to the literature concerning
this type of procedure (is "modular Bayesian" a standard term?).

\spencer{In the revised manuscript we no longer use the modular approach
where some parameters are set as their MLEs before fitting a Bayesian 
model. Instead
we assign a strong prior to the parameter which was giving us trouble
and fit a full Bayesian model. This is
described in the following statement added to revised manuscript:

``In a previous version of this manuscript, when fitting an ASG model
with discrepancy components,
rather than estimating all parameters through posterior updating,
identifiability was encouraged by setting
the parameter $\lambda_s$ from (4)
to be equal to its MLE. In this updated manuscript, 
however, we use empirical Bayes and center the first component of the 
hierarichal prior $\theta$ to be the mean of MLEs for $\lambda_s$ across
all seasons and set the variance parameter so the prior was tight around the
mean. This led to improved mixing of posterior draws and better forecasts.``}

@@

\spencer{Practially, the reason to use the SIR model is generally because it is meant
to capture some of the scientific understanding of a disease outbreak. Again, we argue 
that the ASG is a reasonable alternative to SIR models in the case of the seasonal flu
outbreak. The ASG shape with a rise to a peak and a subsequent fall is similar in shape
to the regular ILI trajectory. It has also been used successfully in forecasting of 
other pheneomena with similar rises and falls (i.e. crop moisture)}.

\item Is the hospitalization indicator specific to influenza? My understanding is that it is, but this
should be more clear in Section 2.2. If it is influenza-specific: to which degree should we expect ILI
to be a helpful predictor given that it also contains COVID, RSV etc.? The influenza dynamics are
likely somewhat hidden in the ILI indicator.

\spencer{Yes, the hospitalization indicator is specific to influenza. See 
response to the next comment (2d.) for how it was addressed in the manuscript.

For your second point on how well ILI can be a helpful predictor of hospitalizations,
we addressed this in the manuscript by adding the following discussion about
figure 6.

``Figure 6 helps
illustrate why this modeling decision was made.
The figure shows scatterplots of ILI percentage and 
hospitalizations on the top row.
The bottom row shows scatterplots of the difference between hospitalizations and 
1 week lags
by ILI percentage at the US national level. The plots in the right column
are of the logarithm of hospitalizations. The lines through the points are
drawn by connecting hospitalizations for consecutive weeks during the flu 
season.
These plots show that a linear or quadratic relationship between ILI percentage 
and hospitalizations does not capture much of the temporal correlation in 
hospitalizations as shown by the obvious loops drawn in the plot. 
However, when including a 
single autoregressive lag of hospitalizations in the model, much of the time 
correlation is
accounted for as shown by the tighter and more frequent crossing of lines
in the bottom row. Models with more than 1 autoregressive
lag showed little or no 
improvement over 1 lag.``}

@@

\item Does the hospitalization indicator refer to incident (new) hospitalization or prevalence, i.e.,
bed occupancy?

\spencer{It refers to new hospitalizations. 
We revised the first paragraph in section 2.2 to read:

``Weekly hospital admission data, which was used as the object of FluSight 
forecasting for the 
2022 and 2023 seasons, is based on the CDC's National Healthcare Safety 
Network (NHSN) dataset entitled \textit{HealthData.gov COVID-19 Reported 
Patient Impact and Hospital Capacity by State Timeseries}. 
In February 2022 it became mandatory for 
all hospitals to report the number of COVID-19 and influenza hospitalizations, 
and since then reporting of hospitalizations has become widespread. These data 
were updated every Wednesday and Friday according to NHSN guidelines 
[22]. Weekly influenza hospitalization data is
defined as the number of newly hospitalized patients with a confirmed diagnosis 
of influenza. From hereon, influenza hospitalizations
will be referrred to as hospitalizations.``}

@@

\item Is there any "mechanistic" rationale behind the ARX formulation for the hospitalizations? If we
predict prevalence / bed occupancy I can sort of see why the differences between weeks would be
explained by ILI. If we predict new hospitalizations I am not sure what motivates this ARX
formulation.

\spencer{ILI is useful as a predictor of hospitalizations, but it fails to 
capture the correlation between consecutive hospitalizations. By adding the 
AR1 component to a model with ILI as a predictor, much of the correlation
structure is then accounted for. 
See the second part of our response to comment 2c. above for how this is 
addressed in the revised manuscript.}

\item Currently the ILI value from the same week is used in equation (7). Would there be a point in
exploring different lags?

\spencer{Exploring different lags is an important idea here, especially considering there
may be a reporting lag. However, the reporting lag is not as big an issue as it has
been in the past. We assessed ILI lags for both one and two weeks previous to the current
week and found that ILI had less predictive power. We note this in the
revised manuscript by saying:

`` We found that models
 with more than 1 autoregressive lag showed 
 little or no forecast improvement over 1 lag.``}

@@

\item On page 14 it says: "all negative values of [...] were set to 0 to reflect realistic values of
hospitalizations". Could this be handled more naturally by appropriate distributional assumptions?

\spencer{We tried both fitting log-hospitalizations as well as assuming 
hospitalizations followed a distribution truncated at 0, but this led to
poorer forecasts or problems fitting the posterior distribution. We make this
statement in the revised manuscript:

``The sample $\{H^*_{s,w + i}\}^K$ was then used as the probabilistic forecast 
for hospitalizations at week $w + i$. 
Occasionally $\{H^*_{s,w + i}\}^K$ included a small number of negative values,
which do not make sense when the distribution is meant to forecast
hospitalizations, a nonnegative number. We attempted to alleviate this problem
by modeling log-hospitalizations or by assuming hospitalizations followed a
distribution truncated at 0, but these models led to poorer forecasts
and issues drawing from the posterior distribution.
For the forecast competition analysis in 
section 5, all negative values of $\{H^*_{s,w + i}\}^K$ were 
set to 0 to reflect realistic values of hospitalizations and comply with the 
FluSight forecasting rules.``}

@@

\item The measure the authors call LWIS tends to give a lot of weight to forecast tasks with low
(expected) incidence as the relative differences between predictions and observations can be quite
large then. The "regular" WIS gives most weight to weeks with high incidence. So in a sense both
are complementary and it may be helpful to also discuss regular WIS a little more.

\spencer{In the revised manuscript, we discuss both the WIS and LWIS in the 
real data analysis section and report both measures. The results for both are
very similar, so to save space we focus a bit more on the LWIS and included 
a little more discussion, including a few plots,
on WIS results in the supplementrary materials.}

\item Why was the CRPS used in the simulation and WIS in the data example? These two are sort of
the same (WIS approximates CRPS), so things would be more coherent if one was used throughout.

\spencer{In the revised manuscript we do use the CRPS, but rather 
use the WIS in both the simulation study
and the real data analysis. We decided on the WIS because that was the measure
used in the FluSight competition.}

@@

\spencer{We updated figures 7 and 9, which show results for the simulation
study and the real data analysis respectively, to include a plot showing
empirical predictive coverage for several nominal levels. This showed that 
our forecasts are reasonably well calibrated relative to others which we 
compare them to.}

\end{enumerate}

\item The authors should at least discuss the possibility of sharing certain parameters across geographic
units as done e.g. in
\href{https://www.nature.com/articles/s41467-021-23234-5}{https://www.nature.com/articles/s41467-021-23234-5}
by Osthus and Moran.

\spencer{The following two statements were added. The first in the main body
and the second in the conslusion.

``Model (1) was fit 
independently for each location. This decision was made largely due to the 
limited time given by FluSight to produce forecasts. That is,
during the competition, data were typically released on Wednesday afternoons and
forecasts were due by the evening of the same day giving only a few hours to 
produce forecasts. A natural extension of the model would be to allow for states
to borrow information from neighboring states.``

``The models in this manuscript were such that the data in each
location were modeled as
independent from all other locations. It is reasonable to assume that a
disease outbreak in one state will share characterstics with the outbreak in 
neighboring states, and indeed the behavior of ILI and hospitalizations is more
closely related for neighboring states as shown in data plots in the
supplementary materials. Notably, the one 2023 FluSight forecast model which 
outperformed
the ASGD\_NORM2, the UGA\_flucast-INFLAenza, was listed as being a spatial
model. And in past seasons, some of the top performing forecast models also 
explicitly modeled spatial relationships [43].``}

@@

Statement at the beginning of section 3.

``Because we desire probabilistic forecasts, Bayesian modeling is a 
natural framework for describing forecast uncertainty. 
Though our main interest is in effeciently producing a probabilistic 
forecast and not 
necessarily in estimating model parameters, certain modeling decisions and 
prior distribution selections proved critical in producing good 
forecasts. After defining the ILI and hospitalization models, 
we describe the 
selection of prior distributions, model fitting, and producing forecasts
from the posterior predictive distribution by combining ILI and hospitalization
models in a way similar to using the 
Bayesian cut [46]."

Here's an example of one of a 
few other statements we added which describes the imporance of 
an informative prior for a 
specific parameter.

``Modeling discrepancy can lead to 
overfitting, particularly in forecasting scenarios, and may also lead to 
identifiability issues. 
Thus, care must be taken in setting parameter constraints as well as in the 
selection of prior distributions [42,8]. 
For instance, when setting $f_{\theta_s}(w)$ to
be the ASG function
in model (1), we found the parameter $\lambda$ from 
in (\ref{eq:asg_function_rep}) difficult to consistently 
estimate precisely because of lack of identifiability. This was resolved by 
assigning to $\lambda$ a more informative prior distribution.``}

@@

``Figure 6 helps
illustrate why this modeling decision was made.
The figure shows scatterplots of ILI percentage and 
hospitalizations on the top row.
The bottom row shows scatterplots of the difference between hospitalizations and 
1 week 
lags
by ILI percentage at the US national level. The plots in the right column
are of the logarithm of hospitalizations. The lines through the points are
drawn by connecting hospitalizations for consecutive weeks during the flu 
season.
These plots show that a linear or quadratic relationship between ILI percentage 
and hospitalizations does not capture much of the temporal correlation in 
hospitalizations as shown by the obvious loops drawn in the plot. 
However, when including a 
single AR lag of hospitalizations in the model, much of the time 
correlation is
accounted for as shown by the tighter and more frequent crossing of lines
in the bottom row. Models with more than 1 AR lag showed little or no 
improvement over 1 lag."}

\item The authors provide also state level predictions. However, I do not see 
where a spatial component is included in the model. Is this only via P or are 
all states just modelled independently. For influenza I would think that it 
might be reasonable to assume that neighbouring states show similar behaviour 
in both ILI and hosptialisation so that borrowing strength between states 
within one model seems sensible. Please discuss.

\spencer{Yes, borrowing strength between states would be reasonable, but 
fitting such a model would likely take much more time than was give by FluSight
to produce forecasts. In the revised manuscript, after defining the ILI model
we state:

``Model (1) was fit 
independently for each location. This decision was made largely due to the 
limited time given by FluSight to produce forecasts. That is,
during the competition, data were typically released on Wednesday afternoons and
forecasts were due by the evening of the same day giving only a few hours to 
produce forecasts. A natural extension of the model would be to allow for states
to borrow information from neighboring states.``

@@

We also added the following to the discussion:

``The models in this manuscript were such that the data in each
location were modeled as
independent from all other locations. It is reasonable to assume that a
disease outbreak in one state will share characterstics with the outbreak in 
neighboring states, and indeed the behavior of ILI and hospitalizations is more
closely related for neighboring states as shown in data plots in the
supplementary materials. Notably, the one 2023 FluSight forecast model which 
outperformed
the ASGD\_NORM2, the UGA\_flucast-INFLAenza, was listed as being a spatial
model. And in past seasons, some of the top performing forecast models also 
explicitly modeled spatial relationships [43].``}

@@

\spencer{In the updated manuscript, this figure was replaced with another figure
which included 3 plots we believe is much more intuitive and descriptive than
the original figure. The title of the new figure remains ``figure 7" and 
the caption is hopefully understandable.}

\item Equation 11, 12, 13, 14: Please motivate. The authors write in Section 3.4 
that the modelling of discrepancy might require careful selection of priors. 
Please underline more how this is taken care of in this manuscript.

\spencer{We decided to put equations 11 - 14 in the supplementary materials, and
give more of an over description of the prior distributions in section 3.6. Here
is one updated paragraph describing how a prior distribution was assigned to
improve identifiability issues introduced by including discprepancy modeling.

``Additional prior constraints were made to improve parameter identifiability. 
Just as was done in Osthus et al. [42],
we set the initial value of the 
susceptible population compartment of the SIR model was set to $S_0 = 0.9$. 
The parameters $I_{0s}$, $\beta_s$, and $\rho_s$ were assigned informative 
priors. In a previous version of this manuscript, when fitting an ASG model
with discrepancy components,
rather than estimating all parameters through posterior updating,
we set the parameter $\lambda_s$ from (4)
to be equal to its MLE. In this updated manuscript, 
however, we use empirical Bayes and center the first component of the 
hierarichal prior $\theta$ to be the mean of MLEs for $\lambda_s$ across
all seasons and set the variance parameter so the prior was tight around the
mean."}

### 2025-08-28 abstract.tex
last revisions

\begin{abstract}
The annual influenza outbreak leads to significant public health and economic 
burdens making it desirable to have prompt and accurate probabilistic forecasts 
of the disease spread. The United States Centers for Disease Control and 
Prevention hosts annually a national flu forecasting competition which has 
led to the development of a variety of flu forecast modeling methods. For the 
first several years of the competition, the target to be forecast was weekly 
percentage of patients with an influenza-like illness (ILI), but in 2021 the 
target was changed to weekly hospitalization counts. Reliable state and national 
hospitalization data has only been available since 2021, but for ILI the data 
has been available since 2010 and has been successfully forecast for several 
seasons.
In this manuscript, we introduce a two component 
Bayesian modeling framework for 
forecasting weekly hospitalizations utilizing both hospitalization data and ILI 
data. The first component is for modeling ILI data using a dynamic nonlinear 
Bayesian hierarchical model. This component includes modeling discrepancy 
between a functional model and the observed which requires thoughtful decisions
for modeling constraints and selecting prior parameter distributions.
The second component is for modeling 
hospitalizations as a function of ILI. For hospitalization forecasts, ILI is 
first forecasted and then hospitalizations forecasts are produced using
ILI forecasts 
as a linear or quadratic predictor. The combination of the models is done in a
manner similar to using the Bayesian cut.
In a simulation study, two ILI forecast 
models, including one similar to the winning model for two seasons of the CDC 
forecast competition from Osthus et al.
% \cite[]{osthus2019dynamic} 
% 
and a nonlinear Bayesian hierarchical model from Ulloa 
% \cite[]{ulloa2019}
are compared with other standard forecasts models.
The usefulness of including a systematic model 
discrepancy term in the ILI model is specifically addressed and proves 
to greatly improve forecasts. In a real data analysis, 
forecasts of state and national 
hospitalizations for the 2023-24 flu season are made and compared with forecasts
from 20 other competing models. The forecasts from the proposed methodology 
outperformed all but one of 20 competing models. 
\end{abstract}

>>>>

\begin{abstract}
The annual influenza outbreak leads to significant public health and economic 
burdens making it desirable to have prompt and accurate probabilistic forecasts 
of the disease spread. The United States Centers for Disease Control and 
Prevention hosts annually a national flu forecasting competition which has 
led to the development of a variety of flu forecast modeling methods. For the 
first several years of the competition, the target to be forecast was weekly 
percentage of patients with an influenza-like illness (ILI), but in 2021 the 
target was changed to weekly hospitalization counts. Reliable state and national 
hospitalization data has only been available since 2021, but for ILI the data 
has been available since 2010 and has been successfully forecast for several 
seasons.
In this manuscript, we introduce a two component 
Bayesian modeling framework for 
forecasting weekly hospitalizations utilizing both hospitalization data and ILI 
data. The first component is for modeling ILI data using a nonlinear 
Bayesian hierarchical model. This component includes modeling discrepancy 
between a nonlinear functional model and the observed ILI which requires 
thoughtful decisions
for modeling constraints and selecting prior parameter distributions.
The second component is for modeling 
hospitalizations as a function of ILI. For hospitalization forecasts, ILI is 
first forecasted and then hospitalizations forecasts are produced using
ILI forecasts 
as a linear or quadratic predictor. The combination of the models is done in a
manner similar to using the Bayesian cut.
In a simulation study, two ILI forecast 
models, including one similar to the winning model for two seasons of the CDC 
forecast competition from Osthus et al.
% \cite[]{osthus2019dynamic} 
% 
and a nonlinear Bayesian hierarchical model from Ulloa 
% \cite[]{ulloa2019}
are compared with other standard forecasts models.
The usefulness of including a systematic model 
discrepancy term in the ILI model is specifically addressed and proves 
to greatly improve forecasts. In a real data analysis, 
forecasts of state and national 
hospitalizations for the 2023-24 flu season are made and compared with forecasts
from 20 other competing models. The forecasts from the proposed methodology 
outperformed all but one of 20 competing models. 
\end{abstract}
